---
title: "Homowork 10-HarleeRamos"
author: "Harlee Ramos"
date: "2024-11-27"
output: html_document
---

<b style="color: green;">Question 2. The dataset LogisticData.csv includes 200 individuals, each described by the following variables:</b><br>

* **Age:** Age of the individual (in years)

* **Income:** Annual income of the individual (in dollars)

* **Education:** Education level of the individual, categorized as:

    * HighSchool
    
    * College
    
    * Graduate
    
* **Purchase:** Binary response variable indicating whether the individual made a purchase:

    * 1 = Yes
    
    * 0 = No
    
<br>
The goal is to build a logistic regression model to predict the probability of purchase (Purchase) based on the predictors.
<br>
(a) Load the dataset into R and inspect how the reference level for Education is set using the code below.<br>
```{r}
# Load the dataset
data <- read.csv("LogisticData.csv")

# Checking how the default reference level is set
data$Education <- factor(data$Education)
levels(data$Education)

# Set 'HighSchool' as the reference level
data$Education <- relevel(data$Education, ref = "HighSchool")
levels(data$Education)
```
As shown in the output, “College” is currently set as the default reference level.
Follow the instructions in the code below to set ’HighSchool’ as the new reference level.
<br>
That is, there are two indicator (dummy) variables such as:<br>
Education_College =
{
1, if the person has a terminal bachelor's degree
0, otherwise
}
<br>
Education_Graduate =
{
1, if the person has a terminal graduate degree
0, otherwise
}
<br>


(b) Fit a logistic regression model using Age, Income, and Education as predictors for Purchase. Write the estimated model equation.
```{r cars}
# Fit the logistic regression model
model1 <- glm(Purchase ~ Age + Income + Education, data = data, family = binomial)

# View the summary of the model
summary(model1)
```


(c) From the Summary, interpret the coefficients of the logistic regression model. Specifically: <br>

(i) How do the log-odds of purchasing change with each additional year of age?<br>
The coefficient for Age is 0.06727. This means that for every additional year of age, the log-odds of purchasing increase by 0.06727, assuming all other variables remain constant.


(ii) How do the log-odds of purchasing change with every additional dollar in crease in income? <br>
The coefficient for Income is 0.00002744. This means that for every additional dollar of income, the log-odds of purchasing increase by 0.00002744, assuming all other variables remain constant.

(iii) Similar to (i) and (ii) above, how would you interpret the coefficients for the Education levels?<br>
The reference category for Education is HighSchool, as indicated by the releveling. The coefficients for College and Graduate represent the change in log-odds compared to HighSchool.<br>
	<br>
	•	EducationCollege (0.2579):
Individuals with a college education have 0.2579 higher log-odds of purchasing compared to those with a high school education, assuming other variables remain constant.<br>
	<br>
	•	EducationGraduate (1.264):
Individuals with a graduate degree have 1.264 higher log-odds of purchasing compared to those with a high school education, assuming other variables remain constant.
<br>

(d) Use the model to predict the probability of purchase for an individual with:<br>
Age = 30,
Income = $50,000,
Education = College.
<br>
The following code may be helpful.

```{r}
# Predict the probability for an individual with Age = 30, Income = 50000, Education = College
new_data <- data.frame(Age = 30, Income = 50000, Education = factor("College", levels = c("HighSchool", "College", "Graduate")))
predicted_prob <- predict(model1, new_data, type = "response")
predicted_prob
```
<br>
(e) Evaluate the model’s performance by creating a confusion matrix. What is the accuracy of the prediction?<br>
The following code may be helpful.<br>

The model has high accuracy (89%), but its performance is skewed toward correctly predicting non-purchases (specificity) while being poor at identifying actual purchases (low sensitivity). 

```{r}
# Load required library
library(caret)

# Generate predictions (threshold = 0.5 for classification)
predictions <- ifelse(predict(model1, type = "response") > 0.5, 1, 0)

# Create confusion matrix
conf_matrix <- confusionMatrix(as.factor(predictions), as.factor(data$Purchase))
conf_matrix
```



<b style="color: green;">Question 3. This question need the data file PoissonData.csv. Please see the context of the data below:</b><br>
Context of the Data A community health organization is studying the number of emergency room (ER) visits for individuals in a city. The dataset includes the following variables:
<br>
* **ER Visits:** Number of ER visits in a year (response variable, count data).
<br>
* **Age:** Age of the individual (in years).
<br>
* **Income:** Annual income of the individual (in $1,000s).
<br>
* **Insurance:** Binary variable indicating whether the individual has health insurance (1 = Yes, 0 = No).
<br>

(a) Plot histograms for the ER Visits variable. What do you observe about its
distribution? The following code may be helpful.<br>
```{r}
# Load the dataset
data <- read.csv("PoissonData.csv")
hist(data$ER_Visits, main="Histogram of ER Visits",
     xlab="Number of ER Visits",col="skyblue",border="black")
```




(b) Fit a Poisson regression model with ER Visits as the response variable and Age, Income, and Insurance as predictors. Write out the equation of the fitted model.
The following code may be helpful.

```{r}
# Fit a Poisson regression model
model2 <- glm(ER_Visits ~ Age + Income + Insurance, 
             family = poisson(link = "log"), 
             data = data)

# View model summary
summary(model2)
```


(c) Interpret the coefficients for each predictor in the model. Specifically:<br>

(i) How will the expected number of ER visits change with a one-year increase in age?<br>
For every additional year of age, the log of the expected number of ER visits decreases by 0.0222, holding all other variables constant.
To interpret this in terms of the expected number of ER visits, exponentiate the coefficient:<br>

Expected ER Visits Ratio (per 1-year increase in Age) = e^(-0.0222) is approx. 0.978.<br>
The expected number of ER visits decreases by about 2.2% for every additional year of age.<br>

(ii) How will the expected number of ER visits change with a $1,000 increase in income?<br>
For every $1,000 increase in income, the log of the expected number of ER visits increases by 0.0322, holding all other variables constant.<br>
Exponentiating the coefficient:<br>

Expected ER Visits Ratio (per $1,000 increase in Income) = e^(0.0322) is approx. 1.033.<br>
The expected number of ER visits increases by about 3.3% for every $1,000 increase in income.<br>

(iii) Based on the model, how does health insurance affect the expected number of ER visits?<br>
Having health insurance (Insurance = 1) reduces the log of the expected number of ER visits by 0.6015, compared to not having insurance (Insurance = 0), holding all other variables constant.<br>
Exponentiating the coefficient:<br>
Expected ER Visits Ratio (with Insurance vs. without Insurance) = e^(-0.6015) is approx.0.548.<br>

Then individuals with health insurance are expected to make about 54.8% as many ER visits as those without insurance, or in other words, having health insurance reduces expected ER visits by 45.2%.

(d) Predict the expected number of ER visits for a 40-year-old individual with an annual income of $50,000 who has health insurance.<br>
The following code may be helpful.<br>

```{r}
# Predict for a 40-year-old with $50,000 income and health insurance
new_data <- data.frame(Age = 40, Income = 50, Insurance = 1)

# Predict expected number of ER visits
predicted_visits <- predict(model2, newdata = new_data, type = "response")

# Print the prediction
cat("Predicted number of ER visits:", predicted_visits, "\n")
```









